---
title: "qmm"
author: "phani varshitha"
date: "2023-09-24"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Objective Function:The Objective function is to Maximize the profit (Z)

Max Z = 420L1 + 360M1 + 300S1 + 420L2 + 360M2 + 300S2 + 420L3 + 360M3 + 300S3

Where P1, P2, P3 are the Plant1, Plant2 and Plant3 and L, M, S be the sizes of the product.
Therefore, Li,Mi,Si will be the sizes in planti where i=1,2,3

Subject to the below constraints:

L1 + M1 + S1 <= 750

L2 + M2 + S2 <= 900

L3 + M3 + S3 <= 450

20L1 + 15M1 + 12S1 <= 13000

20L2 + 15M2 + 12S2 <= 12000

20L3 + 15M3 + 12S3 <= 5000

L1 + L2 + L3 <= 900

M1 + M2 + M3 <= 1200

S1 + S2 + S3 <= 750

The constraints are written as follows

L1 + M1 + S1 + 0L2 + 0M2 + 0S2 + 0L3 + 0M3 + 0S3 <= 750

0L1 + 0M1 + 0S1 + L2 + M2 + S2 + 0L3 + 0M3 + 0S3 <= 900

0L1 + 0M1 + 0S1 + 0L2 + 0M2 + 0S2 + L3 + M3 + S3 <= 450

20L1 + 15M1 + 12S1 + 0L2 + 0M2 + 0S2 + 0L3 + 0M3 + 0S3 <= 13000

0L1 + 0M1 + 0S1 + 20L2 + 15M2 + 12S2 + 0L3 + 0M3 + 0S3 <= 12000

0L1 + 0M1 + 0S1 + 0L2 + 0M2 + 0S2 + 20L3 + 15M3 + 12S3 <= 5000

L1 + 0M1 + 0S1 + L2 + 0M2 + 0S2 + L3 + 0M3 + 0S3 <= 900

0L1 + M1 + 0S1 + 0L2 + M2 + 0S2 + 0L3 + M3 + 0S3 <= 1200

0L1 + 0M1 + S1 + 0L2 + 0M2 + S2 + 0L3 + 0M3 + S3 <= 750

Solving LP Model Using R
```{r}

library(lpSolve)

object_func <- c(420, 420, 420,
                360, 360, 360,
                300, 300, 300)

left_const <- matrix(c(1, 1, 1, 0, 0, 0, 0, 0, 0,
                        0, 0, 0, 1, 1, 1, 0 ,0, 0,
                        0, 0, 0, 0, 0, 0, 1, 1, 1,
                        20, 15, 12, 0, 0, 0, 0, 0, 0,
                        0, 0, 0, 20, 15, 12, 0, 0, 0,
                        0, 0, 0, 0, 0, 0, 20, 15, 12,
                        1, 0, 0, 1, 0, 0, 1, 0, 0,
                        0, 1, 0, 0, 1, 0, 0, 1, 0,
                        0, 0, 1, 0, 0, 1, 0, 0, 1,
                        900, 900, 900, -750, -750, -750, 0, 0, 0,
                        0, 0, 0, 450, 450, 450, -900, -900, -900,
                        450, 450, 450, 0, 0, 0, -750, -750, -750), nrow = 12, byrow = TRUE)

right_const <- c(750,
                 900,
                 450,
                 13000,
                 12000,
                 5000,
                 900,
                 1200,
                 750,
                 0,
                 0,
                 0)

signs <- c("<=",
              "<=",
              "<=",
              "<=",
              "<=",
              "<=",
              "<=",
              "<=",
              "<=",
              "=",
              "=",
              "=")

```
```{r}
lp("max",object_func,left_const,signs,right_const)
```
```{r}
lp("max",object_func,left_const,signs,right_const )$solution
```
