---
title: "QMM_group14"
author: "Dev,Sakshi,Phani"
date: "2023-12-10"
output:
  pdf_document: default
  html_document: default
---

## Problem Statement

You have been assigned the task of forming groups. Consider that your class consists of 15 students, and you aim to create 5 groups, each containing 3 students. The main goal is to maximize the chance of success for each group on a class project.

***

## Solution

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading the required packages

library(tidyverse)
library(ggplot2)
library(dplyr)

```

**Creating a table for the data:**

```{r}
set.seed(69)
df= data.frame("Student ID"=1:15,
               "Student name"=c("Emily", "James", "Isabella", "Ethan", "Sophia", "Alexander", "Olivia", "Michael", "Ava", "William", "Mia", "Benjamin", "Charlotte", "Jacob", "Amelia"),
               GPA=round(runif(15,min=1.5,max = 4),1),
               Attendance=sample(5:10,15,replace = TRUE),
               "Number_of_Awards"=sample(1:5,15,replace = TRUE),check.names = FALSE)
df=df %>% 
  mutate(Points=round(0.4*GPA+0.5*Attendance+0.1*Number_of_Awards,0))
df
```

**Solving the constructed problem:**

```{r}
library(lpSolveAPI)

#Making lp problem 

lp=make.lp(0,75)

#Making objective function 

set.objfn(lp,rep(as.numeric(df$Points),5))

#Setting sense as maximum 

lp.control(lp,sense='max')

#Making type of problem as integer

set.type(lp,75,"integer")
```

**Making constraints**

```{r}

#One group can only have 3 students:

add.constraint(lp, c(rep(1,15),rep(0,60)), "=",3)
add.constraint(lp, c(rep(0,15),rep(1,15),rep(0,45)), "=",3)
add.constraint(lp, c(rep(0,30),rep(1,15),rep(0,30)), "=",3)
add.constraint(lp, c(rep(0,45),rep(1,15),rep(0,15)), "=",3)
add.constraint(lp, c(rep(0,60),rep(1,15)), "=",3)

#One student can only be in one group: 

add.constraint(lp,c(1,rep(0,14),1,rep(0,14),1,rep(0,14),1,rep(0,14),1,rep(0,14)),"=",1)
add.constraint(lp,c(0,1,rep(0,13),0,1,rep(0,13),0,1,rep(0,13),0,1,rep(0,13),0,1,rep(0,13)),'=',1)
add.constraint(lp,c(rep(0,2),1,rep(0,12),rep(0,2),1,rep(0,12),rep(0,2),1,rep(0,12),rep(0,2),1,rep(0,12),rep(0,2),1,rep(0,12)),'=',1)
add.constraint(lp,c(rep(0,3),1,rep(0,11),rep(0,3),1,rep(0,11),rep(0,3),1,rep(0,11),rep(0,3),1,rep(0,11),rep(0,3),1,rep(0,11)),'=',1)
add.constraint(lp,c(rep(0,4),1,rep(0,10),rep(0,4),1,rep(0,10),rep(0,4),1,rep(0,10),rep(0,4),1,rep(0,10),rep(0,4),1,rep(0,10)),'=',1)
add.constraint(lp,c(rep(0,5),1,rep(0,9),rep(0,5),1,rep(0,9),rep(0,5),1,rep(0,9),rep(0,5),1,rep(0,9),rep(0,5),1,rep(0,9)),'=',1)
add.constraint(lp,c(rep(0,6),1,rep(0,8),rep(0,6),1,rep(0,8),rep(0,6),1,rep(0,8),rep(0,6),1,rep(0,8),rep(0,6),1,rep(0,8)),'=',1)
add.constraint(lp,c(rep(0,7),1,rep(0,7),rep(0,7),1,rep(0,7),rep(0,7),1,rep(0,7),rep(0,7),1,rep(0,7),rep(0,7),1,rep(0,7)),'=',1)
add.constraint(lp,c(rep(0,8),1,rep(0,6),rep(0,8),1,rep(0,6),rep(0,8),1,rep(0,6),rep(0,8),1,rep(0,6),rep(0,8),1,rep(0,6)),'=',1)
add.constraint(lp,c(rep(0,9),1,rep(0,5),rep(0,9),1,rep(0,5),rep(0,9),1,rep(0,5),rep(0,9),1,rep(0,5),rep(0,9),1,rep(0,5)),'=',1)
add.constraint(lp,c(rep(0,10),1,rep(0,4),rep(0,10),1,rep(0,4),rep(0,10),1,rep(0,4),rep(0,10),1,rep(0,4),rep(0,10),1,rep(0,4)),'=',1)
add.constraint(lp,c(rep(0,11),1,rep(0,3),rep(0,11),1,rep(0,3),rep(0,11),1,rep(0,3),rep(0,11),1,rep(0,3),rep(0,11),1,rep(0,3)),'=',1)
add.constraint(lp,c(rep(0,12),1,rep(0,2),rep(0,12),1,rep(0,2),rep(0,12),1,rep(0,2),rep(0,12),1,rep(0,2),rep(0,12),1,rep(0,2)),'=',1)
add.constraint(lp,c(rep(0,13),1,0,rep(0,13),1,0,rep(0,13),1,0,rep(0,13),1,0,rep(0,13),1,0),'=',1)
add.constraint(lp,c(rep(0,14),1,rep(0,14),1,rep(0,14),1,rep(0,14),1,rep(0,14),1),'=',1)

#GPA constraint: 

G=as.numeric(df$GPA)
add.constraint(lp, c(G,rep(0,60)), ">=",2.5)
add.constraint(lp, c(rep(0,15),G,rep(0,45)), ">=",2.5)
add.constraint(lp, c(rep(0,30),G,rep(0,30)), ">=",2.5)
add.constraint(lp, c(rep(0,45),G,rep(0,15)), ">=",2.5)
add.constraint(lp, c(rep(0,60),G), ">=",2.5)

#Attendance constraint:

A=as.numeric(df$Attendance)
add.constraint(lp, c(A,rep(0,60)), ">=",7)
add.constraint(lp, c(rep(0,15),A,rep(0,45)), ">=",7)
add.constraint(lp, c(rep(0,30),A,rep(0,30)), ">=",7)
add.constraint(lp, c(rep(0,45),A,rep(0,15)), ">=",7)
add.constraint(lp, c(rep(0,60),A), ">=",7)

#Awards constraint: 

Aw=as.numeric(df$Number_of_Awards)
add.constraint(lp, c(Aw,rep(0,60)), ">=",3)
add.constraint(lp, c(rep(0,15),Aw,rep(0,45)), ">=",3)
add.constraint(lp, c(rep(0,30),Aw,rep(0,30)), ">=",3)
add.constraint(lp, c(rep(0,45),Aw,rep(0,15)), ">=",3)
add.constraint(lp, c(rep(0,60),Aw), ">=",3)

#Setting bounds 

set.bounds(lp,lower = rep(0,75),upper = rep(1,75))
```

```{r}

#Getting zero by solving lp
solve(lp)

# We got 0 as the solution which means that model works absolutely fine and we can further indulge into it to find other parts as well.

#Getting objective function:
P=as.numeric(df$Points)
get.objective(lp)
```

**Getting Points of each group:**

```{r}
#Group-1
sum(get.variables(lp)[1:15]*P)

#Group-2
sum(get.variables(lp)[16:30]*P)

#Group-3
sum(get.variables(lp)[31:45]*P)

#Group-4
sum(get.variables(lp)[46:60]*P)

#Group-5
sum(get.variables(lp)[61:75]*P)

```

**Getting Students in each group:**

```{r}

#Group-1
get.variables(lp)[1:15]

#Group-2
get.variables(lp)[16:30]

#Group-3
get.variables(lp)[31:45]

#Group-4
get.variables(lp)[46:60]

#Group-5
get.variables(lp)[61:75]

```

**Name of students in each group:**

- **Group-1:**

```{r}
df %>% 
  filter(`Student ID`%in% c(1,3,13)) %>% 
  select(`Student name`)
```

- **Group-2:**

```{r}
df %>% 
  filter(`Student ID` %in% c(2,8,11)) %>% 
  select(`Student name`)
```

- **Group-3:**

```{r}
df %>% 
  filter(`Student ID` %in% c(5,7,15)) %>% 
  select(`Student name`)
```

- **Group-4:**

```{r}
df %>% 
  filter(`Student ID`%in% c(4,6,10)) %>% 
  select(`Student name`)
```

- **Group-5:**

```{r}
df %>% 
  filter(`Student ID` %in% c(9,12,14)) %>% 
  select(`Student name`)
```

***
