---
title: "Assignment_3"
author: "phani varshitha"
date: "2023-10-15"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
AED_data = matrix(c(20,14,25,400,100,12,15,14,300,125,10,12,15,500,150,80,90,70,"-","-"),ncol = 5, byrow = TRUE)
colnames(AED_data)= c ("WH1","WH2","WH3","production cost","production capacity")
rownames(AED_data)=c("Plant_A","Plant_B","Plant_c","monthly demand")
AED_data=as.table(AED_data)
AED_data
```
#objective function
$$TC\hspace{0.1cm}=420x_{11}+\hspace{0.1cm}414x_{12}+\hspace{0.1cm}425x_{13}+\hspace{0.1cm}312x_{21}+\hspace{0.1cm}315x_{22}+\hspace{0.1cm}314x_{23}+\hspace{0.1cm}510x_{31}
+\hspace{0.1cm}512x_{32}+\hspace{0.1cm}515x_{33}$$
```{r}
library(lpSolve)
AED_2 = matrix(c(420,414,425,100,"p1",312,315,314,125,"p2",510,512,515,150,"p3",80,90,70,250,"-","q1","q2","q3"),ncol = 5,nrow = 5,byrow = TRUE)
colnames(AED_2)=c("WH1","WH2","WH3","production capacity","dual_supply")
rownames(AED_2)=c("Plant_A","Plant_B","Plant_c","monthly demand","dual-demand")
AED_2= as.table(AED_2)
AED_2
```
# dual constraints 
$$ \hspace{.1cm}X_{11}+X_{12}+X_{13}\le100$$
$$ \hspace{.1cm}X_{21}+X_{22}+X_{23}\le125$$
$$ \hspace{.1cm}X_{31}+X_{32}+X_{33}\le150$$
#3.demand constraint
$$ \hspace{.1cm}X_{11}+X_{11}+X_{11}\ge80$$
$$ \hspace{.1cm}X_{12}+X_{22}+X_{32}\ge90$$
$$ \hspace{.1cm}X_{13}+X_{23}+X_{33}\ge70$$
#non-negativity of variables
$$ \hspace{.1cm}x_{ij}\ge0\hspace{0.1cm}where\hspace{0.1cm}i=1,2,3\hspace{0.1cm}and\hspace{0.1cm}j=1,2,3$$
I took on the transportation issue in the R programming language, which is intrinsically unbalanced because of an imbalance between supply and demand. In the fourth column of the cost matrix, I included a "dummy" variable to account for the case where demand is 10 units short. The dummy variable has a demand of 10 units and a transportation cost of 0. Even if the supply and demand limitations in the mathematical model are originally unbalanced, this modification helps to assure that they are.

```{R}

# AEDs_Costs matrix
AEDs_Costs <- matrix(c(420,414,425,0,
312,315,314,0,
510,512,515,0),ncol = 4,byrow=TRUE)
## Set the names of the rows (constraints) and columns (decision variables)
colnames(AEDs_Costs) <- c("WareHouse1", "WareHouse2", "WareHouse3","Dummy")
rownames(AEDs_Costs) <- c("Plant A", "Plant B", "Plant C")
AEDs_Costs
```
# performing signs and right hand side 
```{r}
row.signs = rep("<=",3)
row.rhs = c(100,125,150)
col.signs = rep(">=",4)
col.rhs = c(80,90,70,10)
```
#solve lp model
```{r}
lptrans = lp.transport(AEDs_Costs,"min",row.signs,row.rhs,col.signs,col.rhs)
```

```{r}
lptrans$solution
```
```{r}
lptrans$objval
```

```{r}
#Objective function
obj <- c(420, 414, 425, 312, 315, 314, 510, 512, 515)

# Coefficients matrix for constraints (lhs)
lhs <- matrix(c(1, 1, 1, 0, 0, 0, 0, 0, 0,
                0, 0, 0, 1, 1, 1, 0, 0, 0,
                0, 0, 0, 0, 0, 0, 1, 1, 1,
                1, 0, 0, 1, 0, 0, 1, 0, 0,
                0, 1, 0, 0, 1, 0, 0, 1, 0,
                0, 0, 1, 0, 0, 1, 0, 0, 1), nrow = 6, byrow = TRUE)

# Right-hand side values for constraints (rhs)
rhs <- c(100, 125, 150, 80, 90, 70)

# Define the direction of inequalities (equalities)
dir <- c("<=", "<=", "<=", "=", "=", "=")

# Solve the linear programming problem
lp=lp("min", obj, lhs, dir, rhs)

# Extract the solution
solution <- lp$solution

# Print the optimal solution
print(solution)

```
**The interpretation of the dual variables** 

For the supply constraints(1,2,3): Dual variables indicate the marginal cost of producing an additional unit of the product at each plant. A higher dual variable value for a plant implies that increasing production capacity at that plant would incur a higher cost compared to the other plants.

For the demand constraints(1,2,3): Dual variables represent the marginal revenue or profit obtained by supplying an additional unit of the product to each warehouse. A higher dual variable value for a warehouse suggests that fulfilling additional demand at that warehouse would yield a higher profit compared to the other warehouses.
